---
title: "Chapter 10 Lab"
author: "Xinyu Li"
date: "February 22, 2017"
output: html_document
---

## 10.4 Lab 1: Principal Components Analysis

```{r}
states = row.names(USArrests)
names(USArrests)
apply(USArrests, 2, mean) # 2 means applying mean function to the columns
apply(USArrests, 2, var)
```

```{r}
pr.out = prcomp(USArrests, scale=TRUE)
names(pr.out)
pr.out$center # mean of variables
pr.out$scale # standard deviations of variables
pr.out$rotation
dim(pr.out$x) # columns of x are principal component scores
biplot(pr.out, scale=0)
# change the sign of principal components
pr.out$rotation = - pr.out$rotation
pr.out$x = - pr.out$x
biplot(pr.out, scale=0)
pr.out$sdev
pr.var = pr.out$sdev^2 # variance explained by each principal component
pr.var
pve = pr.var/sum(pr.var)
pve
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1), type="b")
plot(cumsum(pve), xlab=" Principal Component ", ylab ="Cumulative Proportion of Variance Explained ", ylim=c(0,1), type="b")
```

## 10.5 Lab 2: Clustering

## 10.5.1 K-Means Clustering

```{r}
set.seed(2)
x = matrix(rnorm(50*2), ncol=2)
x[1:25,1] = x[1:25,1] + 3
x[1:25,2] = x[1:25,2] - 4
km.out = kmeans(x, 2, nstart=20)
km.out$cluster
plot(x, col=(km.out$cluster+1), main="K-Means Clustering Results with K=2", xlab="", 
     ylab="", pch=20, cex=2)
```

```{r}
set.seed(4)
km.out = kmeans(x, 3, nstart=20)
km.out
plot(x, col=(km.out$cluster+1), main="K-Means Clustering Results with K=3", xlab="", 
     ylab="", pch=20, cex=2)
```

If a value of nstart greater than one is used, then K-means clustering will be performed using multiple random assignments in Step 1 of Algorithm 10.1, and the kmeans() function will report only the best results.

```{r}
set.seed(3)
km.out = kmeans(x, 3, nstart=1)
km.out$tot.withinss # total within-cluster sum of squares
km.out = kmeans(x, 3, nstart=20)
km.out$tot.withinss
km.out$withinss # individual within-cluster sum-of-squares
```

## 10.5.2 Hierarchical Clustering

```{r}
hc.complete = hclust(dist(x), method="complete")
hc.average = hclust(dist(x), method="average")
hc.single = hclust(dist(x), method="single")
par(mfrow=c(1,3))
plot(hc.complete, main =" Complete Linkage ", xlab="", sub ="", cex =.9)
plot(hc.average, main =" Average Linkage ", xlab="", sub ="", cex =.9)
plot(hc.single, main =" Single Linkage ", xlab="", sub ="", cex =.9)
cutree(hc.complete, 2)
cutree(hc.average, 2)
cutree(hc.single, 2)
cutree(hc.single, 4)
```

```{r}
x = matrix(rnorm(30*3), ncol=3)
dd = as.dist(1-cor(t(x)))
plot(hclust(dd, method="complete"), main=" Complete Linkage with Correlation -Based Distance ", xlab="", sub ="")
xsc = scale(x)
plot(hclust(dist(xsc), method="complete"), main =" Hierarchical Clustering with Scaled Features ")
```

## 10.6 Lab 3: NCI60 Data Example

```{r}
library(ISLR)
nci.labs = NCI60$labs
nci.data = NCI60$data
dim(nci.data)
nci.labs[1:4]
table(nci.labs)
```


## 10.6.1 PCA on the NCI60 Data

```{r}
pr.out = prcomp(nci.data, scale=TRUE)

Cols = function(vec) {
  cols = rainbow(length(unique(vec)))
  return(cols[as.numeric(as.factor(vec))])
}

par(mfrow=c(1,2))
plot(pr.out$x[,1:2], col=Cols(nci.labs), pch=19, xlab="Z1", ylab="Z2")
plot(pr.out$x[,c(1,3)], col=Cols(nci.labs), pch=19, xlab="Z1", ylab="Z3")

summary(pr.out)
pve = 100*pr.out$sdev^2/sum(pr.out$sdev^2)
par(mfrow=c(1,2))
plot(pve, type="o", ylab="PVE", xlab="Principal Component", col="blue")
plot(cumsum(pve), type="o", ylab="Cumulative PVE", xlab="Principal Component", col="brown3")
par(mfrow=c(1,2))
plot(summary(pr.out)$importance[2,], type="o", ylab="PVE", xlab="Principal Component", col="blue")
plot(summary(pr.out)$importance[3,], type="o", ylab="Cumulative PVE", xlab="Principal Component", col="brown3")
```

## 10.6.2 Clustering the Observations of the NCI60 Data

```{r}
sd.data = scale(nci.data)
par(mfrow=c(1,3))
data.dist = dist(sd.data)
plot(hclust(data.dist), labels=nci.labs, main="Complete Linkage", xlab="", sub="", ylab="")
plot(hclust(data.dist, method="average"), labels=nci.labs, main="Average Linkage", xlab="", sub="", ylab="")
plot(hclust(data.dist, method="single"), labels=nci.labs, main="Single Linkage", xlab="", sub="", ylab="")

hc.out = hclust(dist(sd.data))
hc.out
hc.clusters = cutree(hc.out, 4)
table(hc.clusters, nci.labs)

par(mfrow=c(1,1))
plot(hc.out, labels=nci.labs)
abline(h=139, col="red")

set.seed(2)
km.out = kmeans(sd.data, 4, nstart=20)
km.clusters = km.out$cluster
table(km.clusters, hc.clusters)

hc.out = hclust(dist(pr.out$x[,1:5]))
plot(hc.out, labels=nci.labs, main="Hier. Clust . on First Five Score Vectors")
table(cutree(hc.out,4), nci.labs)
```